---
title: "Assignment9"
author: "Ugo Iroh"
date: "2023-01-30"
output: html_document
---

`Using a for loop and an if statement, write a function to calculate the number of zeroes in a numeric vector. Before entering the loop, set up a vector of integers with some zero values. Then, set up a counter variable counter <- 0. Inside the loop, add 1 to counter each time you have a zero in the vector. Finally, use return(counter) for the output.`

```{r}
my_vec <- c(rep(0, 5), rep(3, 5))
print(my_vec)
counter <- 0
for(i in 1:length(my_vec)){
if(my_vec[i]==0) 
  {counter <- counter + 1
}  
}
counter
```


`Write a function that takes as input two integers representing the number of rows and columns in a matrix. The output is a matrix of these dimensions in which each element is the product of the row number x the column number.`
```{r}

my_func <- function(x=10, y=10){
   my_matrix <- matrix(nrow=x, ncol=y)
for (i in 1:nrow(my_matrix)){
  for(j in 1:ncol(my_matrix)){
   my_matrix[i,j] <- i * j
  }
}
  return(my_matrix)
}
 
my_func() #call function
#my_func(x=15,y=5) ##keep running with different values

```


`Use a for loop to repeat your previous functions from Weekly Assignment 2 (which simulate a data frame and extract a p-value) 100 times. Store the results in a data frame that has 1 column indicating the replicate number and 1 column for the p-values (a total of 2 columns and 100 rows).`
```{r}
## for(variable in a sequence){#starts the for loop
#body of the for loop
#end of for loop


##Simulate data frame

myFunc <- function(mean1=10, mean2=45){

  Type <-c(rep("Control",25), rep("Treatment",25))
  Value <- c(rnorm(n=25, mean=mean1, sd=1.5), rnorm(n=25, mean=mean2, sd=2))
  df <- data.frame(Type, Value)
  return(df)
}

myFunc()
#myFunc(mean1=20, mean2=50) ##run with different values


#Extract p-value
my_pval <- function(dataframe=myFunc()){
 df.aov <- aov(Value ~Type, data = dataframe)
 pval <- summary(df.aov)[[1]][["Pr(>F)"]][1]
 return(pval)
  
}
my_pval()  


loop <- data.frame(num_rep=NA, p_val=NA)
for(i in 1:100){
  loop[i,1] <- i
  df <- myFunc()
  pVal <- my_pval(df)
  loop[i,2] <- pVal
}

print(loop)



```




