---
title: "Assignment9"
author: "Ugo Iroh"
date: "2023-01-30"
output: html_document
---

`Using a for loop and an if statement, write a function to calculate the number of zeroes in a numeric vector. Before entering the loop, set up a vector of integers with some zero values. Then, set up a counter variable counter <- 0. Inside the loop, add 1 to counter each time you have a zero in the vector. Finally, use return(counter) for the output.`

```{r}
my_vec <- c(rep(0, 5), rep(3, 5))
print(my_vec)
counter <- 0
for(i in 1:length(my_vec)){
if(my_vec[i]==0) 
  {counter <- counter + 1
}  
}
counter
```


`Write a function that takes as input two integers representing the number of rows and columns in a matrix. The output is a matrix of these dimensions in which each element is the product of the row number x the column number.`
```{r}

my_func <- function(x=10, y=10){
   my_matrix <- matrix(nrow=x, ncol=y)
for (i in 1:nrow(my_matrix)){
  for(j in 1:ncol(my_matrix)){
   my_matrix[i,j] <- i * j
  }
}
  return(my_matrix)
}
 
my_func() #call function
my_func(x=15,y=5) #keep running with different values

```


`Use a for loop to repeat your previous functions from Weekly Assignment 2 (which simulate a data frame and extract a p-value) 100 times. Store the results in a data frame that has 1 column indicating the replicate number and 1 column for the p-values (a total of 2 columns and 100 rows).`
```{r}




```




